---
title: "raster-cubes-tmap"
author: "Martijn Tennekes"
date: "9/18/2020"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation of stars and starsdata 

We recommend to install the github version of **stars**, and the CRAN version of **tmap**:
  
```{r, eval=FALSE}
if (!require(remotes))
install.packages("remotes")
remotes::install_github("r-spatial/stars")
install.packages("starsdata", repos = "https://gis-bigdata.uni-muenster.de/pebesma", type = "source")
install.packages("tmap")
```

Load the packages:
  
```{r}
library(stars)
library(tmap)
```

## Vector data cube example

This example is the North Carolina SIDS (sudden infant death syndrome) data set. See https://edzer.github.io/UseR2020/#Vector_data_cube:_North_Carolina_SIDS
  
The code to create the vector cube is the following:
  
```{r}
L7file = system.file("tif/L7_ETMs.tif", package = "stars")
(L7 = read_stars(L7file))
```



## Plot it with tmap

Star objects can be plot with `tm_raster` as follows:
  
```{r}
tm_shape(L7) +
  tm_raster()
```

Note that most values are around 50 to 100. In order to create more contrast, we can use the *kmeans* algorithm, which creates clusters (by default 5). Note that data for which we apply kmeans is one dimensional.

```{r}
tm_shape(L7) +
  tm_raster(style = "kmeans")
```

We can also create an rgb image as with `tm_rgb`:

```{r}
tm_shape(L7) +
  tm_rgb(1, 2, 3)
```

The image is a bit too dark due to the low values; recall that most values are around 50-100, while the value range is from 0 to 255.
Let us look at the histogram:

```{r}
hist(L7[[1]], breaks = 100)
```

Since `tm_rgb` plot the data values 'as is', so without transformations, we can apply the transformation in advance.


```{r}
library(dplyr)
L7_mod = L7 %>% st_apply(3, pmax, 25) %>% 
  st_apply(3, pmin, 150) %>% 
  '-'(25)

tm_shape(L7_mod) +
  tm_rgb(1, 2, 3, max.value = 125)
```


